---
title: "Attractiveness Ratings Example"
author: "Erin M. Buchanan"
date: "`r Sys.Date()`"
output: word_document
  # html_document:
  #   toc: true
  #   toc_depth: 4
  #   toc_float: true
---

### Vignette Setup:

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

# Set a random seed
set.seed(5989320)

# Libraries necessary for this vignette
library(rio)
library(flextable)
library(dplyr)
library(tidyr)
library(psych)

# Function for simulation
item_power <- function(data, # name of data frame
                       dv_col, # name of DV column as a character
                       item_col, # number of items column as a character
                       sample_start = 20, 
                       sample_stop = 200, 
                       sample_increase = 5,
                       decile = .5){
  
  DF <- cbind.data.frame(
    "dv" = data[ , dv_col],
    "items" = data[ , item_col]
  )
  
  # just in case
  colnames(DF) <- c("dv", "items")
  
  # figure out the "sufficiently narrow" ci value
  SE <- tapply(DF$dv, DF$items, function (x) { sd(x)/sqrt(length(x)) })
  cutoff <- quantile(SE, probs = decile)
  
  # sequence of sample sizes to try
  samplesize_values <- seq(sample_start, sample_stop, sample_increase)

  # create a blank table for us to save the values in 
  sim_table <- matrix(NA, 
                      nrow = length(samplesize_values), 
                      ncol = length(unique(DF$items)))

  # make it a data frame
  sim_table <- as.data.frame(sim_table)

  # add a place for sample size values 
  sim_table$sample_size <- NA

  # loop over sample sizes
  for (i in 1:length(samplesize_values)){
      
    # temp that samples and summarizes
    temp <- DF %>% 
      group_by(items) %>% 
      sample_n(samplesize_values[i], replace = T) %>% 
      summarize(se = sd(dv)/sqrt(length(dv)))
    
    # dv on items
    colnames(sim_table)[1:length(unique(DF$items))] <- temp$items
    sim_table[i, 1:length(unique(DF$items))] <- temp$se
    sim_table[i, "sample_size"] <- samplesize_values[i]
    
  }

  # figure out cut off
  final_sample <- sim_table %>% 
    pivot_longer(cols = -c(sample_size)) %>% 
    rename(item = name, se = value) %>% 
    group_by(sample_size) %>% 
    summarize(Percent_Below = sum(se <= cutoff)/length(unique(DF$items))) 
  
  # multiply by correction 
  final_sample$new_sample <- round(39.369 + 0.700*final_sample$sample_size + 0.003*cutoff - 0.694*length(unique(DF$items)))
  
  return(list(
    SE = SE, 
    cutoff = cutoff, 
    DF = DF, 
    sim_table = sim_table, 
    final_sample = final_sample
  ))

}
```

### Project/Data Title: 

Attractiveness Ratings 

Data provided by: Carlota Batres 

### Project/Data Description: 

This dataset contains 200 participants rating 20 faces on attractiveness. Ethical approval was received from the Franklin and Marshall Institutional Review Board and each participant provided informed consent. All participants were located in the United States. Participants were instructed that they would be viewing several faces which were photographed facing forward, under constant camera and lighting conditions, with neutral expressions, and closed mouths. Each participant would have to rate the attractiveness of the presented faces. More specifically, participants were asked "How attractive is this face?", where 1 = "Not at all attractive" and 7 = "Very attractive". Participants rated each face individually, in random order, and with no time limit. Upon completion, participants were paid for participation in the study.

### Methods Description: 

The data was collected online using Amazon’s Mechanical Turk platform.  

### Data Location: 

Included with the vignette. 

```{r}
DF <- import("batres_data.sav")

str(DF)
```

### Date Published: 

No official publication date.

### Dataset Citation: 

Batres, C. (2022). Attractiveness Ratings. [Data set]. 

### Keywords: 

faces, ratings

### Use License: 

Attribution-NonCommercial-ShareAlike CC BY-NC-SA

### Geographic Description - City/State/Country of Participants:

United States

### Column Metadata: 

```{r}
metadata <- import("batres_metadata.xlsx")

flextable(metadata) %>% autofit()
```

### AIPE Analysis:

The data should be in long format with each rating on one row of data. 

```{r}
# Reformat the data
DF_long <- pivot_longer(DF, cols = -c(Participant_Number)) %>% 
  rename(item = name, score = value)

flextable(head(DF_long)) %>% autofit()
```

#### Stopping Rule

```{r}
# Function for simulation
var1 <- item_power(data = DF_long, # name of data frame
            dv_col = "score", # name of DV column as a character
            item_col = "item", # number of items column as a character
            sample_start = 20, 
            sample_stop = 300, 
            sample_increase = 5,
            decile = .5)
```


What the usual standard error for the data that could be considered for our stopping rule using the 50% decile? 

```{r}
# individual SEs
var1$SE

var1$cutoff
```

Using our 50% decile as a guide, we find that `r round(var1$cutoff, digits = 3)` is our target standard error for an accurately measured item. 

#### Minimum Sample Size

To estimate minimum sample size, we should figure out what number of participants it would take to achieve 80%, 85%, 90%, and 95% of the SEs for items below our critical score of `r round(var1$cutoff, digits = 3)`?

Please note that you will always need to simulate larger than the pilot data sample size to get the starting numbers. We will correct them below. As shown in our manuscript, we need to correct for the overestimation of sample sizes based on the original pilot data size. Given that the pilot data is large: `r nrow(DF)`, this correction is especially useful. This correction is built into our function. 

```{r}
flextable(var1$final_sample %>% 
            filter(Percent_Below >= .80) %>% 
            arrange(Percent_Below, new_sample)) %>% 
  autofit()
```

Our minimum suggested sample size does not exist at exactly 80% of the items, but instead we can use 85% (*n* = `r var1$final_sample %>% filter(Percent_Below >= .80) %>% arrange(Percent_Below, new_sample) %>% slice_head() %>%  pull(new_sample)` as the minimum). In this example, 90% actually hits below 85% at the minimum (*n* = `r var1$final_sample %>% filter(Percent_Below >= .90) %>% arrange(Percent_Below, new_sample) %>% slice_head() %>%  pull(new_sample)`). For 95% of items hitting criterion, the minimum sample size would be approximately `r var1$final_sample %>% filter(Percent_Below >= .95) %>% arrange(Percent_Below, new_sample) %>% slice_head() %>%  pull(new_sample)`. In this case, the researcher could decide that smallest number would be their minimum sample size. 

#### Maximum Sample Size

While there are many considerations for maximum sample size (time, effort, resources), the simulation suggests that `r var1$final_sample %>% filter(Percent_Below >= .95) %>% arrange(Percent_Below, new_sample) %>% slice_head() %>%  pull(new_sample)` people would ensure nearly all items achieve cutoff criterions. 

#### Final Sample Size

In any estimate for sample size, you should also consider the potential for missing data and/or unusable data due to any other exclusion criteria in your study (i.e., attention checks, speeding, getting the answer right, etc.). In this study, we likely expect all participants to see all items and therefore, we could expect to use the minimum sample size as our final sample size, the point at which all items reach our SE criterion, or the maximum sample size. 

