---
title: "Seeing is Believing"
author: "Jason Geller"
date: "Last Knitted: `r Sys.Date()`"
# output: word_document
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

### Vignette Setup: 

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

# Libraries necessary for this vignette
library(rio)
library(flextable)
library(dplyr)
library(tidyr)
library(semanticprimeR)
set.seed(92747)

# Function for simulation
item_power <- function(data, # name of data frame
                       dv_col, # name of DV column as a character
                       item_col, # number of items column as a character
                       sample_start = 20, 
                       sample_stop = 200, 
                       sample_increase = 5,
                       decile = .5){
  
  DF <- cbind.data.frame(
    dv = data[ , dv_col],
    items = data[ , item_col]
  )
  
  # just in case
  colnames(DF) <- c("dv", "items")
  
  # figure out the "sufficiently narrow" ci value
  SE <- tapply(DF$dv, DF$items, function (x) { sd(x)/sqrt(length(x)) })
  cutoff <- quantile(SE, probs = decile)
  
  # sequence of sample sizes to try
  samplesize_values <- seq(sample_start, sample_stop, sample_increase)

  # create a blank table for us to save the values in 
  sim_table <- matrix(NA, 
                      nrow = length(samplesize_values), 
                      ncol = length(unique(DF$items)))

  # make it a data frame
  sim_table <- as.data.frame(sim_table)

  # add a place for sample size values 
  sim_table$sample_size <- NA

  # loop over sample sizes
  for (i in 1:length(samplesize_values)){
      
    # temp that samples and summarizes
    temp <- DF %>% 
      group_by(items) %>% 
      sample_n(samplesize_values[i], replace = T) %>% 
      summarize(se = sd(dv)/sqrt(length(dv)))
    
    # dv on items
    colnames(sim_table)[1:length(unique(DF$items))] <- temp$items
    sim_table[i, 1:length(unique(DF$items))] <- temp$se
    sim_table[i, "sample_size"] <- samplesize_values[i]
    
  }

  # figure out cut off
  final_sample <- sim_table %>% 
    pivot_longer(cols = -c(sample_size)) %>% 
    dplyr::rename(item = name, se = value) %>% 
    group_by(sample_size) %>% 
    summarize(percent_below = sum(se <= cutoff)/length(unique(DF$items))) 
  
  # multiply by correction 
  final_sample$new_sample <- round(39.369 + 0.700*final_sample$sample_size + 0.003*cutoff - 0.694*length(unique(DF$items)))
  
  return(list(
    SE = SE, 
    cutoff = cutoff, 
    DF = DF, 
    sim_table = sim_table, 
    final_sample = final_sample
  ))

}
```

### Project/Data Title: 

Seeing Is Believing: How Media Type Effects Truth Judgements

Data provided by: Gianni Ribeiro

### Project/Data Description: 

People have been duly concerned about how fake news influences the minds of the populous since the rise of propaganda in World War One (Lasswell, 1927). Experts are increasingly worried about the effects of false information spreading over the medium of video. Members of the deep trust alliance, a global network of scholars researching deepfakes and doctored videos, state that ‘a fundamental erosion of trust is already underway’ (Harrison, 2020). Newman et al. (2015) discovered that the media type through which information is presented does indeed affect how true the information feels. Newman speculated that this truthiness effect could be because images provide participants with more information than text alone, thus making the source feel more informationally rich.

In this experiment, our aim is to test the generalizability of Newman’s truthiness effect in two ways: first, to see if it extends to other media types in addition to images, and second, to test if it applies to other domains. In this study, we will present individuals with true and false claims presented through three different media types: (1) text, (2) text alongside a photo, and (3) text alongside a video. This is a direct replication of Newman’s experiment, just with the addition of the video condition. Similarly, participants will also be asked to make truth judgements about trivial claims and claims about COVID-19, to see if the truthiness effect extends to other domains besides trivia.

In this within-subjects design, participants will be presented with true and false claims about trivia and COVID-19 in counterbalanced order. These claims will be randomly assigned to appear either as text alone, text alongside an image, or text alongside a video. Participants will be asked to rate how true they believe each claim is.

### Methods Description: 

Participants were largely sourced from the first-year participant pool at The University of Queensland. Participation was completely voluntary, and participants can choose to withdraw at any time.

Thirty matched trivia claims were generated directly from Newman’s materials. These claims were selected and a true and false version of each claim was created. Newman’s original claims are available at the following link: https://data.mendeley.com/datasets/r68dcdjrpc/1 

The second set of materials comprising of matched true and false claims was generated using information resources from the World Health Organisation, and various conspiracy websites. These claims were then fact-checked by Kirsty Short, an epidemiologist and senior lecturer in the School of Chemistry and Molecular Sciences at The University of Queensland.

The claims were also pilot tested to ensure that none of them performed at floor or ceiling. This pilot test consisted of  56 participants and subsequently four claims were dropped. The data from this pilot test was also used to accurately perform a power analysis. After generating the means of the pilot test, we found that to acquire a power of 0.8 or greater, there must be a mean difference of 0.4 between each media type. This mean difference is quite conservative, since we plan to measure truth ratings on a six-point scale and is easily achievable with 100 participants.

The videos were largely sourced from the stock image website Envato Elements and Screenflow’s Royalty Free Stock Media Library.

### Data Location: 

Data can be found here: https://osf.io/zu9pg/
 
```{r}
DF <- import("moat_data.csv") 
  
str(DF)
```

### Date Published: 

No official publication, see citation below.  

### Dataset Citation: 

Moat, K., Tangen, J., & Newman, E. (2021). Seeing Is Believing: How Media Type Effects Truth Judgements.

### Keywords: 

covid, trivia games, claims, interpretation

### Use License: 

Open access with reference to original paper (Attribution-NonCommercial-ShareAlike CC BY-NC-SA)

### Geographic Description - City/State/Country of Participants:

Brisbane, Queensland, Australia

### Column Metadata:

```{r}
metadata <- tibble::tribble(
             ~Variable.Name,                                                                                                       ~Variable.Description, ~`Type.(numeric,.character,.logical,.etc.)`,
                      "Id",                                                                                                            "Participant ID",                                   "numeric",
                   "Domain",                                                    "Whether the trial is a claim about COVID ('covid') or TRIVIA ('trivia)",                                 "character",
                   "Medium", "Whether the trial appears as text alone ('claim'), text alongside an image ('photo'), or text alongside a video ('video')",                                 "character",
               "Trial_type",                                        "Whether the trial presents a claim that is TRUE ('target') or FALSE ('distractor')",                                 "character",
                   "Rating",                           "Paritcipant’s truth rating of the claim ranging from 1 (definitely false) to 6 (definitely tue)",                                   "numeric"
             )

flextable(metadata) %>% autofit()
```

### AIPE Analysis:

```{r}
# Function for simulation
var1 <- item_power(data = DF, # name of data frame
            dv_col = "rating", # name of DV column as a character
            item_col = "question_type", # number of items column as a character
            sample_start = 20, 
            sample_stop = 300, 
            sample_increase = 5,
            decile = .4)
```

#### Stopping Rule

What is the usual standard error for the data that could be considered for our stopping rule?


```{r}
var1$SE
var1$cutoff

cutoff <- var1$cutoff

# we can also use semanticprimer's function
cutoff_score <- calculate_cutoff(population = DF,
                                 grouping_items = "question_type",
                                 score = "rating",
                                 minimum = min(DF$rating),
                                 maximum = max(DF$rating))
cutoff_score$cutoff
```

Using our 40% decile as a guide, we find that `r round(cutoff, digits = 3)` is our target standard error for an accurately measured item. 

#### Minimum Sample Size

To estimate minimum sample size, we should figure out what number of participants it would take to achieve 80%, 85%, 90%, and 95% of the SEs for items below our critical score of `r round(cutoff, digits = 3)`?

How large does the sample have to be for 80% of the items to be below our stopping SE rule? 

```{r}
flextable(var1$final_sample %>% head()) %>% 
  autofit()

final_table <- calculate_correction(
  proportion_summary = var1$final_sample,
  pilot_sample_size = DF %>% group_by(question_type) %>% 
    summarize(sample_size = n()) %>% ungroup() %>% 
    summarize(avg_sample = mean(sample_size)) %>% pull(avg_sample),
  proportion_variability = cutoff_score$prop_var
  )

flextable(final_table) %>% 
  autofit()
```

Based on these simulations, we can decide our minimum sample size is likely close to `r final_table %>% arrange(percent_below, corrected_sample_size) %>% filter(percent_below >= 80) %>% slice_head() %>% pull(corrected_sample_size) %>% round()`.

#### Maximum Sample Size

In this example, we could set our maximum sample size for 90% of items, which would equate to `r final_table %>% arrange(percent_below, corrected_sample_size) %>% filter(percent_below >= 90) %>% slice_head() %>% pull(corrected_sample_size) %>% round()` participants. 

#### Final Sample Size

You should also consider any potential for missing data and/or unusable data given the requirements for your study. Given that participants are likely to see all items in this study, we could use the minimum, stopping rule, and maximum defined above. 


