---
title: "Superficial face judgement "
author: "Necdet Gürkan & Yangyang Yu"
date: "`r Sys.Date()`"
output: word_document
  # html_document:
  #   toc: true
  #   toc_depth: 4
  #   toc_float: true
---

### Vignette Setup:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Libraries necessary for this vignette
library(rio)
library(flextable)
library(dplyr)
library(tidyr)
library(psych)
```

### Project/Data Title: 

Deep models of superficial trait inferences

Data provided by: Jordan W. Suchow

### Project/Data Description: 

The diversity of human faces and the contexts in which they appear gives rise to an expansive stimulus space over which people infer psychological traits (e.g., trustworthiness or alertness) and other attributes (e.g., age or adiposity). Machine learning methods, in particular deep neural networks, provide expressive feature representations of face stimuli, but the correspondence between these representations and various human attribute inferences is difficult to determine because the former are high-dimensional vectors produced via black box optimization algorithms. In this paper, we combined deep generative image models with over 1 million judgments to model inferences of more than 30 attributes over a comprehensive latent face space. The predictive accuracy of the model approached human interrater reliability, which simulations suggest would not have been possible with fewer faces, fewer judgments, or lower-dimensional feature representations. The model can be used to predict and manipulate inferences with respect to arbitrary face photographs or to generate synthetic photorealistic face stimuli that evoke impressions tuned along the modeled attributes.

In sum, the dataset contains 1.14 million ratings across 1000 items and 34 traits by 5,000 participants. NOTE: The trait trustworthy in the dataset is collected twice, so the trait column has 35 traits.


### Methods Description: 

For the attribute model studies, we used a between-subjects design where participants evaluated faces with respect to each attribute. Participants first consented. Then they completed a preinstruction agreement to answer open-ended questions at the end of the study. In the instructions, participants were given 25 examples of face images in order to provide a sense of the diversity they would encounter during the experiment. Participants were instructed to rate a series of faces on a continuous slider scale where extremes were bipolar descriptors such as “trustworthy” to “not trustworthy.” We did not supply definitions of each attribute to participants and instead relied on participants’ intuitive notions for each. 

Each participant then completed 120 trials with the single attribute to which they were assigned. One hundred of these trials displayed images randomly selected (without replacement) from the full set; the remaining 20 trials were repeats of earlier trials, selected randomly from the 100 unique trials, which we used to assess intrarater reliability. Each stimulus in the full set was judged by at least 30 unique participants. 

At the end of the experiment, participants were given a survey that queried what participants believed we were assessing and asked for a self-assessment of their performance and feedback on any potential points of confusion, as well as demographic information such as age, race, and gender. Participants were given 30 min to complete the entire experiment, but most completed it in under 20 min. Each participant was paid $1.50. 

### Data Location: 

https://github.com/jcpeterson/omi

```{r}
## Please set the work directory to the folder containing the scripts and data
face_data <- import("suchow_data.csv")
str(face_data)
```

### Date Published:

2022-04-15

### Dataset Citation:

Peterson, J. C., Uddenberg, S., Griffiths, T., Todorov, A., & Suchow, J. W. (2022). Deep models of superficial face judgments. Proceedings of the National Academy of Sciences (PNAS).

### Keywords: 

first impressions, social perception, face perception

### Use License: 

https://creativecommons.org/licenses/by-nc-sa/4.0/

### Geographic Description - City/State/Country of Participants:

For the attribute model studies, we used Amazon Mechanical Turk to recruit a total of 4,157 participants across 10,974 sessions, of which 10,633 (≈ 97%) met our criteria for inclusion. Participants identified their gender as female (2,065) or male (2,053), preferred not to say (21), or did not have their gender listed as an option (18). The mean age was ∼39 y old. Participants identified their race/ethnicity as either White (2,935), Black/African American (458), Latinx/a/o or Hispanic (158), East Asian (174), Southeast Asian (71), South Asian (70), Native American/American Indian (31), Middle Eastern (12), Native Hawaiian or Other Pacific Islander (3), or some combination of two or more races/ethnicities (215). The remaining participants either preferred not to say (22) or did not have their race/ethnicity listed as an option (8). Participants were recruited from the United States.

### Column Metadata: 

```{r}
metadata <- import("suchow_metadata.xlsx")

flextable(metadata) %>% autofit()
```

### AIPE Analysis: 

#### Stopping Rule

When pilot data is this large, it is important to sample a smaller subset based on what the participant might actually do in the study. We will pick 50 faces rated on 10 traits - and then select the highest and lowest variance to estimate from. This choice is somewhat arbitrary - in a real study, you could chose to use only the variables you were interested in and pick the most conservative values or simply average together estimates from all variables. 

```{r}
# pick random faces
faces <- unique(face_data$stimulus)[sample(unique(face_data$stimulus), size = 50)]
# pick random traits
traits <- unique(face_data$trait)[sample(unique(face_data$trait), size = 10)]

face_data <- face_data %>% 
  filter(trait %in% traits) %>% 
  filter(stimulus %in% faces)
```

```{r sd analysis}
# all SEs 
SE_full <- tapply(face_data$response, face_data$trait, function (x) { sd(x)/sqrt(length(x)) })
SE_full
```

```{r subset and restructure}
## smallest variance is trait 4
face_data_trait4_sub <- subset(face_data, trait == names(which.min(SE_full)))

## largest is trait 30
face_data_trait30_sub <- subset(face_data, trait == names(which.max(SE_full)))
```

```{r compute se for traits}
# individual SEs for 4 trait 
SE1 <- tapply(face_data_trait4_sub$response, face_data_trait4_sub$stimulus, function (x) { sd(x)/sqrt(length(x)) })
quantile(SE1, probs = .5)

# individual SEs for 30 trait
SE2 <- tapply(face_data_trait30_sub$response, face_data_trait30_sub$stimulus, function (x) { sd(x)/sqrt(length(x)) })

quantile(SE2, probs = .5)
```

#### Minimum Sample Size

How large does the sample have to be for 80% and 95% of the items to be below our stopping SE rule? 

```{r power Two different traits}
# sequence of sample sizes to try
samplesize_values <- seq(25, 100, 5)

# create a blank table for us to save the values in 
sim_table <- matrix(NA, 
                    nrow = length(samplesize_values), 
                    ncol = length(unique(face_data_trait4_sub$stimulus)))
# make it a data frame
sim_table <- as.data.frame(sim_table)

# add a place for sample size values 
sim_table$sample_size <- NA
sim_table$var <- "response"

# make a second table for the second variable
sim_table2 <- matrix(NA, 
                    nrow = length(samplesize_values), 
                    ncol = length(unique(face_data_trait30_sub$stimulus)))

# make it a data frame
sim_table2 <- as.data.frame(sim_table2)

# add a place for sample size values 
sim_table2$sample_size <- NA
sim_table2$var <- "response"

# loop over sample sizes for age and outdoor trait
for (i in 1:length(samplesize_values)){
    
  # temp dataframe for age trait that samples and summarizes
  temp7 <- face_data_trait4_sub %>% 
    dplyr::group_by(stimulus) %>% 
    dplyr::sample_n(samplesize_values[i], replace = T) %>% 
    dplyr::summarize(se1 = sd(response)/sqrt(length(response))) 
  
  # 
  colnames(sim_table)[1:length(unique(face_data_trait4_sub$stimulus))] <- temp7$stimulus
  sim_table[i, 1:length(unique(face_data_trait4_sub$stimulus))] <- temp7$se1
  sim_table[i, "sample_size"] <- samplesize_values[i]
  
  # temp dataframe for outdoor trait that samples and summarizes
  temp35 <-face_data_trait30_sub %>% 
    dplyr::group_by(stimulus) %>% 
    dplyr::sample_n(samplesize_values[i], replace = T) %>% 
    dplyr::summarize(se2 = sd(response)/sqrt(length(response))) 
  
  # 
  colnames(sim_table2)[1:length(unique(face_data_trait30_sub$stimulus))] <- temp35$stimulus
  sim_table2[i, 1:length(unique(face_data_trait30_sub$stimulus))] <- temp35$se2
  sim_table2[i, "sample_size"] <- samplesize_values[i]

}
```

```{r summary analysis part1}
cutoff <- quantile(SE1, probs = .5)
final_sample <- 
  sim_table %>%
  pivot_longer(cols = -c(sample_size, var))  %>% 
  dplyr::rename(item = name, se = value)   %>% 
  dplyr::group_by(sample_size, var)  %>% 
  dplyr::summarize(Percent_Below = sum(se <= cutoff)/length(unique(face_data_trait4_sub$stimulus)))  %>% 
  dplyr::filter(Percent_Below >= .80) %>% 
  dplyr::arrange(Percent_Below) %>% 
  mutate(new_sample = round(39.369 + 0.700*sample_size + 0.003*cutoff - 0.694*length(unique(face_data_trait4_sub$stimulus))))

flextable(final_sample) %>% autofit()
```

```{r summary analysis part2}
cutoff <- quantile(SE2, probs = .5) 
final_sample2 <- 
  sim_table2 %>%
  pivot_longer(cols = -c(sample_size, var)) %>% 
  dplyr::rename(item = name, se = value)  %>% 
  dplyr::group_by(sample_size, var) %>% 
  dplyr::summarize(Percent_Below = sum(se <= cutoff)/length(unique(face_data_trait30_sub$stimulus))) %>% 
  dplyr::filter(Percent_Below >= .80) %>% 
  dplyr::arrange(Percent_Below) %>% 
  mutate(new_sample = round(39.369 + 0.700*sample_size + 0.003*cutoff - 0.694*length(unique(face_data_trait4_sub$stimulus))))

flextable(final_sample2) %>% autofit()
```

Based on these simulations, we can decide our minimum sample size is somewhere between `r min(final_sample$new_sample)` and `r min(final_sample2$new_sample)`. We can consider only the most variant trait for power analysis since it would satisfy other traits in the dataset as well.

#### Maximum Sample Size

We could set our maximum sample size for 95% power, which would equate to `r final_sample2 %>% filter(Percent_Below >= .95) %>% pull(new_sample) %>% min()` participants. 

#### Final Sample Size

Considering both estimated traits and a smaller proportion of things to rate, we found that we could likely use samples from 50-75 people. Other considerations could include fatigue on the number of ratings each person has to complete. 